Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2024-11-16 _tex/index.tex
Built site for gh-pages

Surprisingly, despite its reliance on the largest number of simplifying
assumptions \citep[pp.~253]{Bramley_2008}, Case V remains the most
widely used case in the CJ literature. This popularity is largely due to
its simplified statistical representation in the Bradley-Terry-Luce
(BTL) model \citep{Bradley_et_al_1952, Luce_1959}. The BTL model mirrors
Case V's assumptions, with one key difference: while Case V assumes a
Normal distribution for the stimuli's discriminal processes, the BTL
model uses the more mathematically tractable Logistic distribution
\citep[pp.~254]{Bramley_2008} (see Table~\ref{tbl-thurstone_cases}).
This substitution has little effect on the model's estimation or
interpretation, as the Normal and Logistic distributions differ by a
scaling factor of approximately \(1.7\)
\citep[pp.~16]{vanderLinden_et_al_2017_I} (refer to
Figure~\ref{fig-logistic_vs_normal}).

@@

However, Case V (and its BTL model counterpart) was originally developed
to provide a ``rather coarse scaling'' of traits
\citep[pp.~269]{Thurstone_1927b}, prioritizing statistical simplicity
over precision in trait measurement. As a result, its assumptions may
not suit applications beyond the psychophysical contexts for which it
was created. Thurstone himself cautioned that its use ``should not be
made without (an) experimental test'' \citep[pp.~270]{Thurstone_1927b},
acknowledging that some assumptions could prove problematic with more
complex, less homogeneous stimuli, such as handwriting or English
compositions \citep[pp.~374]{Thurstone_1927a}. Consequently, given that
current CJ applications often deal with complex, less homogeneous
stimuli, two key assumptions of Case V may not consistently hold in
theory or practice: the zero correlation and equal dispersion between
stimuli.

>>>>

Surprisingly, despite its reliance on the largest number of simplifying
assumptions
\citetext{\citealp[pp.~253]{Bramley_2008}; \citealp[pp.~677]{Kelly_et_al_2022}},
Case V remains the most widely used case in the CJ literature. This
popularity is largely due to its simplified statistical representation
in the Bradley-Terry-Luce (BTL) model
\citep{Bradley_et_al_1952, Luce_1959}. The BTL model mirrors Case V's
assumptions, with one key difference: while Case V assumes a Normal
distribution for the stimuli's discriminal processes, the BTL model uses
the more mathematically tractable Logistic distribution
\citep[pp.~254]{Andrich_1978, Bramley_2008} (see
Table~\ref{tbl-thurstone_cases}). This substitution has little effect on
the model's estimation or interpretation, as the Normal and Logistic
distributions differ by a scaling factor of approximately \(1.7\)
\citep[pp.~16]{vanderLinden_et_al_2017_I} (refer to
Figure~\ref{fig-logistic_vs_normal}).

@@

However, Case V (and its BTL model counterpart) was originally developed
to provide a ``rather coarse scaling'' of traits
\citep[pp.~269]{Thurstone_1927b}, prioritizing statistical simplicity
over precision in trait measurement \citep[pp.~677]{Kelly_et_al_2022}.
As a result, its assumptions may not suit applications beyond the
psychophysical contexts for which it was created. Thurstone himself
cautioned that its use ``should not be made without (an) experimental
test'' \citep[pp.~270]{Thurstone_1927b}, acknowledging that some
assumptions could prove problematic with more complex, less homogeneous
stimuli, such as handwriting or English compositions
\citep[pp.~374]{Thurstone_1927a}. Consequently, given that current CJ
applications often deal with complex, less homogeneous stimuli, two key
assumptions of Case V may not consistently hold in theory or practice:
the zero correlation and equal dispersion between stimuli.
